---
title: "UDAND_R-project"
author: G.A.C.
date: 2019-02-19
output: html_document
---

```{r setup, include=FALSE}
# suppress code and warnings,
# 
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align='center',
                      fig.path='Figs/', tidy=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
```

The dataset analyzed contains information about the chemical properties of
a sample of almost 5000 white wines, with 11 different measurements and a
'quality' score for each observation. I will attempt to answer the following
question:

##Which chemical properties make a white wine stand out?

###[Skip to Final Plots and Summary](#id1)

***

```{r loading-data-libraries, warning=FALSE, message=FALSE}
#### Loading data and necessary libraries
library(ggplot2)
library(ggpubr) # for arranging plots together
library(GGally) # for a correlation matrix
library(memisc) # for a linear model
wwine <- read.csv('wineQualityWhites.csv')
```

#### A quick look at the data

```{r quick-look}
print('NA values:')
wwine[is.na(wwine)]
print('duplicate observations:')
dim(wwine[duplicated(wwine$X)])[2]
print('structure of the data:')
str(wwine)
```

It looks like there are no missing or duplicate values.
The variable *X* identifies each sample of wine.
The variable *quality* is the dependent variable that I will try to
explain with all the other variables showing chemical properties of each wine.

***

###Univariate analysis
I want to see a summary of the *quality* variable first:

```{r quality-summary}
summary(wwine$quality)
table(wwine$quality)
```
```{r quality-barplot}
# since the variable is not continuous I make a bar plot instead of a histogram
ggplot(wwine, aes(x=factor(quality),
                  color=I('dark green'),
                  fill=I('light green'))) +
  geom_bar() + theme(aspect.ratio = .5)
```

The distribution is 'normal' shaped; the ratings are evenly distributed with
the mean value being very close to the median.


Next I will look at the distribution of the chemical properties that a person
without any deep knowledge about wine or chemistry might feel more familiar
with, looking at the most common values and the variablity in this sample:
*alcohol* content (measured in percentage of volume) and *residual sugar* 
(measured in
[grams per litre](https://en.wikipedia.org/wiki/Sweetness_of_wine#Residual_sugar))
.

I also want to look at the values of *density* and *pH* compared to those of
water (at room temperature,
around [0.998 grams per mililitre](https://en.wikipedia.org/wiki/Density#Water)
and a [pH of 7.0](https://en.wikipedia.org/wiki/PH)).


I will use blue color to mark the inter-quartile range for each variable:
the middle 50% of values between the bottom 25% and the top 25% of all
observations ordered by said variable.

```{r variable_histograms-alcohol}
ggplot(wwine, aes(x=alcohol, color=I('dark blue'),
                  fill=I(ifelse(wwine$alcohol<=quantile(wwine$alcohol, .75),
            ifelse(wwine$alcohol>=quantile(wwine$alcohol,
                                          .25),
                   'light blue', 'grey'), 'grey')))) +
  geom_histogram(binwidth=.1) + ggtitle('Alcohol content, all observations')

summary(wwine$alcohol)
```

The middle 50% of the wines in the sample have an alcohol content between 9.5%
and 11.4%.
The minimum and maximum values are 8.0% and 14.2% and the distribution is
fairly even although there is somewhat of a right skew with the highest values
ranging further past the third quartile than the lowest values extend between
the first quartile and the minimum.


```{r variable_histograms-sugar}
ggplot(wwine, aes(x=residual.sugar, color=I('dark blue'),
                  fill=I(ifelse(wwine$residual.sugar<=
                                  quantile(wwine$residual.sugar, .75),
            ifelse(wwine$residual.sugar>=quantile(wwine$residual.sugar,
                                          .25),
                   'light blue', 'grey'), 'grey')))) +
  geom_histogram(binwidth=.25) +
  ggtitle('Residual Sugar, x axis limited to 99.9th percentile') +
  xlim(c(min(wwine$residual.sugar), quantile(wwine$residual.sugar, .999)))

summary(wwine$residual.sugar)
```

In the case of residual sugar the range of values found in the sample is much
wider and the distribution is very skewed to the right.
The bulk of the distribution ranges from concentrations of about one gram per
litre to values over ten times higher, and there are ouliers up to more than
60 grams per litre that make an x axis limit necessary.
Before seeing any correlation, the high variability in this chemical property
suggests that I should not expect it to have a big impact on the quality of a
white wine.


```{r variable_histograms-density}
ggplot(wwine, aes(x=density, color=I('dark blue'),
                  fill=I(ifelse(wwine$density<=quantile(wwine$density, .75),
            ifelse(wwine$density>=quantile(wwine$density,
                                          .25),
                   'light blue', 'grey'), 'grey')))) +
  geom_histogram(binwidth=0.0002) +
  ggtitle('Density, x axis limited to 99.9th percentile') +
  xlim(c(min(wwine$density), quantile(wwine$density, .999)))

summary(wwine$density)
```

The density of white wine is very similar to that of water and there is little
variability in the values found. There seem to be some outliers but the
distribution is otherwise normal shaped.


```{r variable_histograms-ph}
ggplot(wwine, aes(x=pH, color=I('dark blue'),
                  fill=I(ifelse(wwine$pH<=quantile(wwine$pH, .75),
            ifelse(wwine$pH>=quantile(wwine$pH, .25),
                   'light blue', 'grey'), 'grey')))) +
  geom_histogram(binwidth=.01) +
  ggtitle('pH, all observations')

summary(wwine$pH)
```

White wine is very acidic compared to water and its pH follows a normal
distribution that doesn't appear to indicate big variability.
Still, considering the logarithmic nature of pH it seems that wines at
one end of the distribution might be
much more acidic than those at the other end, so I would need to look at
the correlation with quality to really know if pH is determinant.

***

I decide to make three bins for the *quality* variable and see if I find any
patterns that can answer the question raised earlier:

**Which chemical properties make a white wine stand out?**

#####Response variable bins
First I decide the *quality* scores to divide at, and check that the amount of
observations in each bin is suitable for my purposes:

```{r quality-bins}
wwine$quality.bin <- cut(wwine$quality,
                         breaks=c(0,5,6,9),
                         labels=c('<=5: below average',
                                  '6: average',
                                  '>=7: above average'),
                         include.lowest = TRUE,
                         ordered_result = TRUE)

table(wwine$quality.bin)
```

Then I set a color scale and make a barplot of the new *quality* bins:

```{r qualbin-distribution}
# it will help to set a color scale for the quality bins
qualbin.scale <- scale_fill_manual('Quality bins:', values=c('<=5: below average'='red',
                                       '6: average'='greenyellow',
                                       '>=7: above average'='cyan'))

# I can then add the scale to plots as a layer:
ggplot(wwine, aes(x=quality.bin, fill=quality.bin,
                  color=I('black'))) +
  geom_bar(show.legend=FALSE) +
  qualbin.scale
```

The dividing values I set for the quality bins split the observations
fairly evenly. This is important for any conclusions to be meaningful, as
I do not look for common features of only a tiny minority of the observations.

***

###Bivariate analysis

####Finding the most relevant variables

In order to focus on the more interesting chemical properties, I want to look
at the correlations with *quality*. That way I can choose only a few variables
to analyze further.

```{r correl-heatmap}
ggcorr(wwine[, c(2:13)],
       label=TRUE, label_alpha=TRUE, label_round=2,
       layout.exp=3, hjust=.9, size=4)
```

The results are **not promising: the strongest correlation with *quality***
**is in the variable *alcohol*, with a correlation coefficient of 0.44**.
**Of the other variables** with correlation coefficients with *quality* at
least **close to $\pm0.2$, only *volatile acidity* does not show a stronger**
**correlation of the same sign with *alcohol***.

As I ventured earlier, the highly variable *residual sugar* has a very low
correlation with quality. The effect of *pH* is equally low.

***

####Explanatory variable Box Plots
I make box plots divided by quality bin for each of the explanatory
variables that have a correlation coefficient with *quality*
of at least close to $\pm0.2$: *alcohol* (.44), *volatile acidity* (-.19),
*density* (-.31), *chlorides* (-.21) and *total sulfur dioxide* (-.17).

```{r three-boxplots}
# I assign plots to variables to later arrange them comfortably
al.bp <- ggplot(data=wwine,
       aes(y=alcohol,
           x=quality.bin,
           fill=quality.bin), xaxt='n') +
  geom_boxplot() +
  qualbin.scale +
  ggtitle('Alcohol')

va.bp <- ggplot(data =wwine,
       aes(y=volatile.acidity,
           x=quality.bin,
           fill=quality.bin), xaxt='n') +
  geom_boxplot() +
  qualbin.scale +
  ggtitle('Volatile acidity')

sd.bp <- ggplot(data=wwine,
       aes(y=total.sulfur.dioxide,
           x=quality.bin,
           fill=quality.bin), xaxt='n') +
  geom_boxplot() +
  qualbin.scale +
  ggtitle('Total sulfur dioxide')

# the function ggarrange lets me easily set a common legend for several plots
ggarrange(al.bp, va.bp, sd.bp, ncol=3, common.legend = TRUE, legend='top')
```

Not all the visible trends are strong, however it is possible to see in these
plots how, for example, of the wines rated 'below average', less than 25% (only
those sitting above the higher bound of the IQR -the red box in this case)
reach or exceed the same *alcohol* value that is surpassed by more than 75% of
the wines rated 'above average' (since the lower end of the blue box -the
first quartile of 'above average' wines- lies above the same *alcohol* value).


One can look at the median values and IQRs of the different quality bins for
each chemical property to get an idea of the variable's influence on quality.


#####Dealing with outliers
When outliers are present, limiting the axes to a certain percentile,
for example leaving out the highest 1% of values for the variable,
can help zoom in on the patterns shown by the plot.
If I were to instead subset the data used for the plot, that would cause
slight changes in the measures of median, IQR and so on.

```{r boxplots-with-outliers, echo=FALSE, message=FALSE, warning=FALSE}
dn.bp <- ggplot(data=wwine,
       aes(y=density,
           x=quality.bin,
           fill=quality.bin), xaxt='n') +
  geom_boxplot() +
  qualbin.scale

# I can directly add layers to a plot by its assigned variable name
dn.bp1 <- dn.bp +
  ggtitle('Density, all observations')

dn.bp2 <- dn.bp +
  scale_y_continuous(limits=c(min(wwine$density),
                              quantile(wwine$density, .99))) +
  ggtitle('Density; y limit 99th percentile')

ch.bp <- ggplot(data=wwine,
       aes(y=chlorides,
           x=quality.bin,
           fill=quality.bin)) +
  geom_boxplot(show.legend=FALSE) +
  qualbin.scale

ch.bp1 <- ch.bp +
  ggtitle('Chlorides, all observations')

ch.bp2 <- ch.bp +
  scale_y_continuous(limits=c(quantile(wwine$chlorides, .025),
                              quantile(wwine$chlorides, .975))) +
  ggtitle('Chlorides; y limits 2.5% and 97.5%')

ggarrange(dn.bp1, dn.bp2, ch.bp1, ch.bp2,
          ncol=2, nrow=2, common.legend=TRUE, legend='top')
```


***

####Explanatory variable Histograms

```{r five-histograms, echo=FALSE, message=FALSE, warning=FALSE}
alc.hist <- ggplot(data=wwine, aes(x=alcohol, fill=quality.bin)) +
  geom_histogram(binwidth=.1, show.legend=FALSE) +
  qualbin.scale +
  ggtitle('Alcohol;\nall observations')

tsf.hist <- ggplot(data=wwine, aes(x=total.sulfur.dioxide, fill=quality.bin)) +
  geom_histogram(binwidth=5, show.legend=FALSE) +
  xlim(c(min(wwine$total.sulfur.dioxide),
         quantile(wwine$total.sulfur.dioxide, .99))) +
  qualbin.scale +
  ggtitle('Tot. Sulfur Dioxide;\nx limit 99%')

vla.hist <- ggplot(data=wwine, aes(x=volatile.acidity, fill=quality.bin)) +
  geom_histogram(binwidth=.01, show.legend=FALSE) +
  qualbin.scale +
  xlim(c(min(wwine$volatile.acidity),
         quantile(wwine$volatile.acidity, .975))) +
  ggtitle('Volat. Acidity;\nx limit 97.5%')

den.hist <- ggplot(data=wwine, aes(x=density, fill=quality.bin)) +
  geom_histogram(binwidth=.0002, show.legend=FALSE) +
  qualbin.scale +
  xlim(c(min(wwine$density), quantile(wwine$density, .99))) +
  ggtitle('Density;\nx limit 99%')

chl.hist <- ggplot(data=wwine, aes(x=chlorides, fill=quality.bin)) +
  geom_histogram(binwidth=.001) +
  qualbin.scale +
  xlim(c(min(wwine$chlorides), quantile(wwine$chlorides, .975))) +
  ggtitle('Chlorides;\nx limit 97.5%')

rsg.hist <- ggplot(data=wwine, aes(x=residual.sugar, fill=quality.bin)) +
  geom_histogram(binwidth=.1) +
  qualbin.scale +
  xlim(c(min(wwine$residual.sugar), quantile(wwine$residual.sugar, .99))) +
  ylim(c(0, 100)) +
  ggtitle('Residual Sugar;\nxlim 99%, ylim 100')

ggarrange(alc.hist, vla.hist, den.hist, chl.hist, tsf.hist, rsg.hist,
          ncol=3, nrow=2, common.legend = TRUE, legend='right')
```

Similarly to the box plots, these histograms hint that past a certain threshold
of each of these variables there seems to be a higher likelihood of the wine in
question being 'below average', instead of 'average' or above in quality. It is
difficult to make the same observation about the residual sugar histogram on
the bottom right corner.

#####Explanatory variable bins

It may be interesting to bin each of these variables for more clarity in
later plots.
I write a function **exvr.bin** using the function **cut** to bin values
as low, medium or high dividing them at the 33rd and 67th percentiles, so that
I can more easily bin all the explanatory variables:

```{r explvar-bins, echo=TRUE}
exvr.bin = function(x) cut(x,
                           c(min(x), quantile(x, .333),
                             quantile(x, .667), max(x)),
                         include.lowest=TRUE,
                         labels=c('Bottom 33%',
                                  'Middle 33%',
                                  'Top 33%'),
                         ordered_result=TRUE)

#I can then pipe each column to my function and save time and possible typos:
wwine$alcohol.bin <- wwine$alcohol%>%exvr.bin()
```

```{r apply-bins-scales}
wwine$vol.acid.bin <- wwine$volatile.acidity%>%exvr.bin()
wwine$tot.sulf.bin <- wwine$total.sulfur.dioxide%>%exvr.bin()
wwine$density.bin <- wwine$density%>%exvr.bin()
wwine$chloride.bin <- wwine$chlorides%>%exvr.bin()

# I also set custom scales for the bins I will use:
exvarbin.shape.scale <- scale_shape_manual(name='Chemical property bins:',
                                     values=c('Bottom 33%'=16,
                                              'Middle 33%'=15,
                                              'Top 33%'=17))

exvarbin.fill.scale <- scale_fill_manual('Chemical property bins:',
                                         values=c('Bottom 33%'='navy',
                                                   'Middle 33%'='orange',
                                                   'Top 33%'='magenta'))

exvarbin.color.scale <- scale_color_manual('Chemical property bins:',
                                         values=c('Bottom 33%'='navy',
                                                   'Middle 33%'='orange',
                                                   'Top 33%'='magenta'))

# Beside the previous 'fill' scale for quality bin, I will need a 'color' scale
qualbin.col.scale <- scale_color_manual('Quality',
                                        values=c('<=5: below average'='red',
                                       '6: average'='greenyellow',
                                       '>=7: above average'='cyan'))
```

***



###Multivariate analysis
The truly interesting patterns should become visible when combining several
variables.


####Linear Model
I will first try to make a linear model to more clearly see which variables
are useful for explaining *quality*.


<a id="id2"></a>

First I combine alcohol (which has the strongest correlation with quality) with
the other variables separately.

```{r linear-model1}
m1 <- lm(quality ~ alcohol, data=wwine)
m2 <- update(m1, ~. + volatile.acidity)
m3 <- update(m1, ~. + density)
m4 <- update(m1, ~. + total.sulfur.dioxide)
m5 <- update(m1, ~. + chlorides)
mtable(m1, m2, m3, m4, m5)
```

From the R-squared values I see that the only truly useful variables are
*alcohol* and [*volatile acidity*](https://extension.psu.edu/volatile-acidity-in-wine).

```{r linear-model2}
m6 <- update(m2, ~. + density)
m7 <- update(m2, ~. + total.sulfur.dioxide)
m8 <- update(m2, ~. + chlorides)
m9 <- update(m6, ~. + total.sulfur.dioxide)
m10 <- update(m6, ~. + total.sulfur.dioxide + chlorides)
mtable(m6, m7, m8, m9, m10)
```

This second attempt tells me that keeping *density* might have some value, but I
will most likely discard total sulfur dioxide and *chlorides* because adding
them would hardly improve the model (meaning they would only add noise to a
visualization).


####Scatterplots
With the information from the R-squared values of the linear model I can now
test whether the variables I chose actually add clarity to a visualization.

I want to see scatterplots of the different variables against Quality, and I
can add layers for the bins I made as color and shape. More variables could be
added as size of the points, but this is likely to worsen
overplotting especially given the non-continuous Quality variable and the low
correlations that will not make for clearly separate groups but rather a messy
mix of whatever distinct levels I set.

Since color differences will be the easiest to distinguish, following the
results of
the linear model I will **compare the effect of adding to Alcohol (as x) and**
**Volatile Acidity (as shape) a layer of color determined by the properties of**
**Density, Chlorides or Total Sulfur Dioxide**. To maximize clarity I will use
bins
instead of the original continuous values for all explanatory variables other
than Alcohol.

Note: jitter has been added to the points to mitigate the overplotting caused
by the discrete quality scores. Because of this jitter effect, the points
corresponding to observations in the intersection of a certain quality score
and alcohol content will not appear in exactly the same positions in the y axis
in the different scatterplots. 

```{r scatterplots1}
ggplot(data = wwine,
       aes(y=quality,
           x=alcohol,
           color=density.bin,
           shape=vol.acid.bin)) +
  geom_point(position = 'jitter', alpha=.35) +
  exvarbin.shape.scale + exvarbin.color.scale +
  guides(color=guide_legend(title='Density',
                             override.aes = list(alpha=1)),
         shape=guide_legend(title ='Volatile Acidity')) +
  labs(title='Quality vs Alcohol and Volatile Acidity;\nDensity as color')
```

The distribution of the points hints the positive correlation
between alcohol and quality.
I can tell from this first plot that *density* is strongly inversely correlated
with alcohol, but appears to have very little effect of its own on quality.
If the shape layer reveals anything, it is that high quality wines with
high volatile acidity are almost all high in alcohol.

```{r scatterplots2}
ggplot(data = wwine,
       aes(y=quality,
           x=alcohol,
           color=tot.sulf.bin,
           shape=vol.acid.bin)) +
  geom_point(position = 'jitter', alpha=.35) +
  exvarbin.shape.scale + exvarbin.color.scale +
  guides(color=guide_legend(title='Tot. Sulfur Dioxide',
                             override.aes = list(alpha=1)),
         shape=guide_legend(title ='Volatile Acidity')) +
  labs(title=
    'Quality vs Alcohol and Volatile Acidity;\nTot. Sulfur Dioxide as color')
```

Adding *total sulfur dioxide* as a color layer only shows that it is
correlated with
alcohol, just in a weaker way than density. This variable is of no use in
explaining quality.

```{r scatterplots3}
ggplot(data = wwine,
       aes(y=quality,
           x=alcohol,
           color=chloride.bin,
           shape=vol.acid.bin)) +
  geom_point(position = 'jitter', alpha=.35) +
  exvarbin.shape.scale + exvarbin.color.scale +
  guides(color=guide_legend(title='Chlorides',
                             override.aes = list(alpha=1)),
         shape=guide_legend(title ='Volatile Acidity')) +
  labs(title='Quality vs Alcohol and Volatile Acidity;\nChlorides as color')
```

Similarly to the previous plot, a negative correlation between Alcohol and
Chlorides is apparent, however the three levels of the Chlorides color layer
can be found in similar proportions in every level of Quality. This variable
does not add much value to the visualization.


After seeing the linear model's R-squared values and the color layers in these
scatterplots, I decide to leave out the three variables and focus only on
*alcohol* and *volatile acidity* for my final plots and conclusions.


***
***

<a id="id1"></a> 

##Final Plots and Summary
The attempt to make a linear model makes me decide to focus on the the two
variables *alcohol* and *volatile.acidity* as explanatory variables
for *quality* of the white wine observations, as they seem to be the only ones
from which some meaning can be reliably extracted.



####**Observation counts at each quality score colored by bins of main variables**
The bar plots below show **how many of the observed white wines fall into**
**each discrete quality score**, and then look at this distribution
separately to show **what relative values of the most relevant chemical**
**properties** (bottom, middle and top third of measured values for each)
**make up each quality score**.


```{r final-barplots}
q.alc.bar <- ggplot(data=wwine, aes(x=factor(quality), fill=alcohol.bin)) +
  geom_bar() + exvarbin.fill.scale + ylab(NULL) + xlab(NULL) +
  ggtitle('Colored by\nAlcohol content') + theme_dark()

q.vac.bar <- ggplot(data=wwine, aes(x=factor(quality), fill=vol.acid.bin)) +
  geom_bar() + exvarbin.fill.scale + ylab(NULL) + xlab(NULL) +
  ggtitle('Colored by\nVolatile Acidity') + theme_dark()

final.barplot <- ggarrange(q.alc.bar, q.vac.bar, ncol=2,
                           common.legend=TRUE, legend='top')

annotate_figure(final.barplot,
                top=text_grob(
                  'White wine observation counts at each Quality score
colored by bins of main chemical properties',
                  face='bold', size=15),
                left=text_grob('Observation counts', size=11, face='bold',
                               rot=90),
                bottom=text_grob('Quality scores', size=11, face='bold'))
```


Besides showing what scores are most common, the bar plots show that most wines
rated 5 or worse are relatively low in alcohol while those rated 7 or better
are mostly high in alcohol content, or that relatively low volatile acidity is
much less frequent in wines rated 5 or worse than in better rated ones.


***

####**Distribution of main explanatory variables divided by quality bin**
The following box plots illustrate the most relevant chemical properties'
correlations with the observed white wines' *quality*: positive in the case of
*alcohol* and negative for *volatile.acidity*. The box plots
**divide the white wine observations by quality (average, above, or below)**
and **make it easy to visualize, for each of the chemical properties**, what
the median value is (the value in **the center of the distribution**) and what
values lie within the inter-quartile range (the middle 50% of values between
the lowest 25% and the highest 75%, making **the bulk of the distribution**).
Due to the presence of outliers, one of the plots leaves out the highest 1% of
values to zoom in on the shape of the distribution.


```{r final-boxplots}
# I can use the plots I made earlier by variable name and add layers to them
alc.final.bx <- al.bp + ylab('Alcohol (% of volume)') + xlab(NULL) +
  labs(title='Alcohol content', subtitle='all observations') +
  rremove('x.text')

vac.final.bx <- va.bp + ylab('Volatile Acidity (g/L)') + xlab(NULL) +
  scale_y_continuous(limits=c(min(wwine$volatile.acidity),
                              quantile(wwine$volatile.acidity, .99))) +
  labs(title='Volatile acidity', subtitle='y limit: 99th percentile') +
  rremove('x.text')

final.boxplot <- ggarrange(alc.final.bx, vac.final.bx,
                           common.legend = TRUE, legend='bottom', ncol=2)

annotate_figure(final.boxplot, top=text_grob(
  'Distribution of most relevant chemical properties divided by Quality bin',
  face='bold', size=15))
```


The box plots show graphically how only about a quarter of white wines rated 7
or above in Quality have an alcohol content under 11%, while the
majority of 'average' wines and most of the 'below average' wines have a lower
alcohol content.
Similarly, a value of Volatile Acidity above 0.3 grams per litre makes a white
wine more likely to be rated 'below average', although in the case of this
chemical property wines of quality 'average' and 'above average' do not seem
to differ from each other and there are ouliers in all quality groups;
variability in Volatile Acidity is more acceptable than in Alcohol.


***

####**Quality score explained by main variables**
The scatterplot below can help in visualizing the patterns identified throghout
the analysis.

I plot the continuous values of Alcohol in the x axis because this variable has
the strongest correlation with quality and thus shows the most clear shape.

I represent the three levels of binned values of Volatile Acidity with a color
scale because this is the only other variable with some effect on quality that
is not correlated with Alcohol.

I leave out other variables given their very small influence on quality (as
measured and visualized in the [Multivariate Analysis section](#id2))
and because
their potential to distract the viewer from other patterns is greater
than the value of the information that they could add to the plot.

```{r final-scatterplot, fig.height=6, fig.width=9}
final.scatterplot <- ggplot(data=wwine, 
                              aes(y=factor(quality),
                                  x=alcohol,
                                  color=vol.acid.bin)) +
    geom_point(position='jitter', alpha=.35) +
    exvarbin.color.scale +
    guides(colour=guide_legend(override.aes=list(alpha=1, size=3),
                               title='Volatile Acidity')) +
    theme(aspect.ratio=1) +
    labs(title=
  'Quality score (y) by Alcohol Content (x) and Volatile Acidity (color)',
         subtitle='transparency and jitter added to mitigate overplotting',
         y='Quality score',
         x='Alcohol (% of volume)')
  
  annotate_figure(final.scatterplot,
                  top=text_grob(
       "White wines' quality score explained by main chemical properties",
                    face='bold', size=15))
```

Despite the low correlations in the data, this plot can help see
that the 'above average' wines with low alcohol content have low acidity in
most cases, and most of the highly rated wines with high acidity are also high
in alcohol content.

White wines rated 'average' are a mixed bag, however among those rated 5 or
worse a low volatile acidity is clearly less common.

While some patterns are identifiable, the lack of more clear correlations
suggests that other variables not measured in the available data play an
important role in determining the quality score of a particular white wine.


***


##Reflections
While a conclusion can be extracted that **the most reliable predictors for**
**a white wine being rated highly** (or at least not being rated poorly)
**are first a relatively high alcohol content, and second a**
**relatively low volatile acidity**, there are many observations in the data
that do not match these conditions.

This suggests that **other variables that have not been accounted for** in the
available data **may have an important impact on the quality rating**. Relevant
data other than the chemical measurements available could be the
temperature of each wine when rated or the time since production.


The main difficulty in performing this analysis was the lack of significant
correlations between the response variable and the available explanatory
variables.

Attempting to make a linear model provided valuable insight into which of the
variables were of less relevance, in a way that may be more clear than just
looking at the correlations and plots.

Given these conditions, binning the different variables and then combining them
in multivariate visualizations was useful for seeing the patters present.
Perhaps more detailed binning of the variables and more zoomed-in
visualizations to limit the noise could help extract more precise conclusions.

